---
title: "Exercise_1"
author: "Rachel Kaufman"
date: "2023-08-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("dplyr")
library(dplyr)
```

### **Assignment 1 Stats**
*Questions for Andres are italicized* 

#### are for excersizes
### notes/practice areas following along Andres's explanations :p

#### 1.1
```{r}
typeof(mtcars)
class(mtcars)
```

```{r}
nrow(mtcars) ##number of rows in the data set
ncol(mtcars) #number of columns
length(mtcars) #"length" of the vector, i.e. sa im assuming this is length in terms of longview/long data idk how to describe but its the number of elements within a vector.A dataframe cannot be a data frame unless all the vectors are of the same length. 
dim(mtcars) ##dimensions of the data set
rownames(mtcars) #naming the rows of the data set, aka the main unit of analysis so in this instance all the types of cars that are being assess
colnames(mtcars) ## names of the columns of car data being collected

```

#### 1.2 
```{r}
T <- 123
TRUE <- 123 #error in 
```
Doesn't allow for assigning an object as `TRUE,` this is likely bc TRUE/FALSE are actual commands that you employ on the left side of our <- (*equation? function? what is the right word?*). When assigning T, it isn't taken as true but taken as the character T. 

#### 1.3
when creating atomic vectors, you can coerce them even with different types of datapoints (numeric)
```{r}
lgl <- c(TRUE, FALSE, NA)
int <- c(1L, 6L, NA, 10L)
dbl <- c(1, NA, 2.5, 4.5)
chr <- c(NA, "these are", "some strings")

```
`c()` is for concatenate, this means stringing a list together via combining them. 
so when you implicitly coerce your vectors if they have different types, what that vector will become follows this order: logical --> integer --> double --> character 
For example,
```{r}
str(c(TRUE, "chr"))
dplyr::glimpse(c(TRUE, "chr"))

dplyr:: glimpse(c(117, "my own example!"))
```

Testing my guessing (not really guesssing but *thinking*)
```{r}
c(1, FALSE) #double/number
c("a", 1) #character
c(TRUE, 1L) #integer

dplyr :: glimpse(c(1, FALSE))
dplyr :: glimpse(c("a", 1))
dplyr::glimpse(c(TRUE, 1L))
```

#### 1.4
```{r}
as.logical(T)
X <- TRUE
as.integer(X)
```
this is using the functions to explicitly coerce the value to a new type. SooOo, if it is as.integer, it will be changed to an integer.


```{r}
x <- sample(c(TRUE, FALSE), size = 75, replace = TRUE)

dplyr::glimpse(x)
sum(x)
mean(x)

sum(x)/length(x) #this is literally what mean(x) does bc vector length is the total of the sample/observation count in this silly little example!
```


#### 1.5 
```{r}
## the proportion of cars in the dataset with more than 3 carburators
mean(mtcars$carb > 3)


```

### 1.1.2 subsettting

```{r}

mtcars$mpg

x <- list(chr, lgl, letters)
?list()
dplyr:: glimpse(x)

names(x) <- c("chr", "lgl", "alphabet")
str(x)

```

#### 1.6
so `[]` would be basically saying within the object, like the dataframe mtcars, if you look at ["mpg"] within the mtcars, you would "zoom in" on the subset of two vectors, which is still a data frame. But if you did [["mpg"]] then you are looking within the 2 vectors, which ends up being just the mpg numbers. It's the subsetting of the subsetting

```{r}
class(mtcars["mpg"]) ## includes both car type mpg
class(mtcars[["mpg"]]) ##two brackets leads to 1 
identical(mtcars["mpg"], mtcars$mpg)
identical(mtcars[["mpg"]], mtcars$mpg)
```

#### 1.7
okay, i subsetted the 17th value through the first command, so! woohoo. There is one vector, so no need for a , or anything. 
```{r}
letters[17]

x2 <- seq(from = 2, to = 26, by = 2) #sequence checkkkk
letters[x2]

letters[8:12]
```
letters[[8:12]] --- produces nada! bc not really anything there to subset what has alreqady been subsetted

#### 1.8
```{r}
LETTERS[18] <- NA ##the correct way

letters[letters[18]] <- NA ## the fiddling i didnt understand

letters[letters == "r"] <- NA ## the fiddling i didn't understand pt. 2
 

```
*okay, this i have a question with. by doing it in the two ways i attempted (after doing it the first way via* `LETTERS[18] <- NA` *), it adds on a r NA at the end of the list. Does this effect the type of the object letters? How do I do this without changing the type from it being a vector? Because when I run* `View(letters)` *it now says null. why is this the case? why does the list add on an r and an NA?*  

### 1.1.4 Data Frames
```{r}
mtcars[1:2]
mtcars[,1:3] ## this gives 
mtcars[1:3,] ##this gives us rows

##logical subsetting examples
```

#### 1.9

#### 1.10
#### 1.11
#### 1.2


