---
title: "assignment_week4"
format: html
editor: visual
date: "2023-09-18"
author: "Rachel Kaufman"
mainfont: "Georgia"
format: 
  html: 
    theme: cosmo
    toc: true
    embed-resources: true
  pdf:
    colorlinks: true
execute: 
  echo: false
bibliography: references.bib
csl: american-journal-of-sociology.csl
---

## Assignment 4

```{r}
library(tidyverse)
theme_set(theme_light())
```

#### 4.1

```{r}
x <- rnorm(100000, mean = 0, sd = 1)


result <- as_tibble(x) |> 
  mutate(
    one_sd = ifelse(abs(x) <= 1, 1, 0),
    two_sd = ifelse(abs(x) <= 2, 1, 0),
    above_mean = ifelse(x >= 0, 1, 0),
    below_mean = ifelse(x <= 0, 1, 0)
  )  |> 
  summarize(
    prob_one_sd = mean(one_sd),
    prob_two_sd = mean(two_sd),
    prob_above = mean(above_mean),
    prob_below = mean(below_mean)
  )

result
```

googled for the `abs()` function because i was trying to figure out how to calculate distance from the mean and obviously if i am looking at distance (because mean is at 0 meaning some of my rando numbers will be negative, but i do not want those negatives to be additive, absolute values let me focus on the distance)

#### 4.2

using quantile function

```{r}
quantile(x)
```

#### 4.3

creating specific percentiles

```{r}
quantile(x, probs = c(.005, .995))
```

#### 4.4

Here ya go...

```{r}
as_tibble(x) |> 
 summarize(probability = mean(
   x >= -2.576 & x <= 2.576))
```

#### 4.5

Okay, your equation showed up funny at first but I figured to put it in my viewer thing and it fixed it.

```{r}
sum(runif(n = 20, min = 0, max = 1))

d4.5 <- tibble(id = 1:1e3) |> 
  rowwise() |> 
  mutate(x = sum(runif(20)))  


ggplot(data = d4.5,
       mapping = aes(x)) +
   geom_histogram(bins = 50,
                  color = "darkblue",
                  fill = "lightgray")
```

#### 4.6

Redoing steves simulation!

```{r}
svy_size <- 2247  # number of people in each "poll"

std_error <- sqrt((1/3) * (2/3) / svy_size)
std_error

est_prop <- 1/3

0.33 + 1.96*std_error
0.33 - 1.96*std_error ## "1.96" is roughly "2"!

sd(runif(20)) * sqrt(20)

```
